---
title: "Simulation Example"
date: "Sept 19, 2023"
author: "Samantha-Jo Caetano"
output:
  pdf_document: default
---

# Simulation

## Sample Function

First let's get started with the `sample` function.

```{r}
library(tidyverse)

help(sample) # or ?sample

colour <- c("Red", "Blue", "Green", "Yellow", "Purple")

sample(x=colour, size=3)

```

Let's try using the `set.seed()` function. 

```{r}
sample(x=colour, size=3)
sample(x=colour, size=3)
sample(x=colour, size=3)
sample(x=colour, size=3)

set.seed(4)
sample(x=colour, size=3)

set.seed(4)
sample(x=colour, size=3)

set.seed(4)
sample(x=colour, size=3)
```

Now, let's take a sample with replacement, of size 10, and let's say we know the population have 60% red and 10% of each other colour.

```{r}
set.seed(4)
sample(x=colour, size=10, replace=TRUE, prob=c(0.6, 0.1, 0.1, 0.1, 0.1))

set.seed(4)
sample(x=colour, size=10, replace=TRUE, prob=c(12, 2, 2 ,2, 2))

```

## Simulating from distributions

Similar to `qnorm()` and `pnorm()` one can use `rnorm()` (or `r<well_known_distribution_name>()`) to simulate from a well known distribution.

Exercise: Assume we know waiting times for a bus are usually distributed to be uniform of -1 to +4 minutes (i.e., one minute early to 4 minutes late) and you wait for the bus 10 times in a month. Simulate your ten waiting times.

```{r}
waiting_times <- runif(n=10, min=-1, max=4)

waiting_times
```


## Creating a Tibble

Let's build a dataset/tibble that is about my experiences waiting for the bus. It will include how early/late the bus was (waiting_time), the colour of my shirt the day I was waiting  and the number of people also waiting at the bus stop.

```{r}
set.seed(2023)

waiting_times <- runif(n=10, min=-1, max=4)

shirt_colour <- sample(x=colour, size=10, replace=TRUE, prob=c(0.6, 0.1, 0.1, 0.1, 0.1))

num_peeps_bus <- rpois(n=10, lambda = 3)

my_bus_data <- tibble(waiting_times, shirt_colour, num_peeps_bus)

glimpse(my_bus_data)
```


Let's just take a look at some of the data:

```{r}
my_bus_data %>% ggplot(aes(x=num_peeps_bus)) + geom_histogram()

my_bus_data %>% ggplot(aes(x=waiting_times)) + geom_histogram()
```

Note: There is some code in the Week 1 Module to help you in formatting Tables and Plots in your pdfs.


## Correlating Variables in a Tibble

What if instead you know that waiting_times and num_peeps_bus is correlated?

```{r}
waiting_times <- runif(n=10, min=-1, max=4)

#hist(waiting_times)

shirt_colour <- sample(x=colour, size=10, replace=TRUE, prob=c(0.6, 0.1, 0.1, 0.1, 0.1))

num_peeps_bus <- round(waiting_times + rnorm(n=10, 1, 1))

#hist(num_peeps_bus)

my_bus_data <- tibble(waiting_times, shirt_colour, num_peeps_bus)

```




